---
title: 'EpistasisRank and Epistasis-Katz: incorporating prior knowledge for pathway enrichment and classification'
author: "Saeid Parvandeh"
date: "9/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
setwd("~/Documents/PhD papers/PK/subsampling")
```

## Pathway Analysis

```{r}
library(igraph)
library(SDMTools)
library(Rinbix)
library(ggplot2)
library(caret)
library(xgboost)
library(gridExtra)
library(privateEC)
```

Loading required data for pathway analysis
```{r}
source("helper_functions.R")
# load MDD microarray data
load("camb_fltr.expr.RData")
load("camb_sub_class.RData")
load("jap_fltr.expr.RData")
load("jap_class.RData")

# load reGAIN and IMP query network
load("camb_reGAIN.RData")
load("jap_reGAIN.RData")
load("impNetwork.RData")
```

### Centrality methods without prior knowledge
#### PageRank
```{r}
camb_expre.cor <- cor(camb_fltr.expr)
# co-expression to binary
camb_Adj <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
diag(camb_Adj) <- 0

# Adjacency matrix to graph
camb_Adj_g <- igraph::graph.adjacency(camb_Adj)
# page_woPKRank centrality
camb_page_woPK <- page.rank(camb_Adj_g)$vector
camb_page_woPK.sort <- sort(camb_page_woPK, T)

top_camb_page_woPK <- camb_page_woPK.sort[1:200]
```

#### Katz
```{r}
a <- eigen(camb_Adj)
beta <- rep(1, nrow(camb_Adj))/nrow(camb_Adj)
alpha <- 1/max(a$values) - (1/max(a$values))/100
camb_katz_woPK <- katz.centrality(camb_Adj, alpha, beta)
names(camb_katz_woPK) <- colnames(camb_Adj)
camb_katz_woPK.sort <- sort(camb_katz_woPK, T)

top_camb_katz_woPK <- camb_katz_woPK.sort[1:200]
```

#### EpistasisKatz
```{r}
alpha <- 1/mean(colSums(camb_reGAIN))
beta <- diag(camb_reGAIN)
camb_regain.Katz <- camb_reGAIN
diag(camb_regain.Katz) <- 0
camb_EK_woPK <- katz.centrality(camb_regain.Katz, alpha, beta)
names(camb_EK_woPK) <- rownames(camb_reGAIN) 
camb_EK_woPK.sort <- sort(camb_EK_woPK, T)

top_camb_EK_woPK <- camb_EK_woPK.sort[1:200]
```

#### EpistasisRank
```{r}
camb_ER_woPK <- EpistasisRank(camb_reGAIN, Gamma_vec = .85)
top_camb_ER_woPK <- camb_ER_woPK[1:200, ]
```

#### Coefficient of Variation (cov) filtering
```{r}
exprData <- t(camb_fltr.expr)
top_camb_cov <- t(cov.filter(exprData,.4874)$fdata) # 200 genes
```

### Centrality method with prior knowledge
#### PageRank
```{r}
# prior knowledge
impgraph <- graph_from_adjacency_matrix(impNetwork, mode = "undirected", diag = TRUE, add.rownames = NA)
IMP_deg <- degree(impgraph, normalized = TRUE, loops = FALSE)
camb_page_wPK <- as.numeric(my.pagerank(camb_Adj, IMP_deg))
names(camb_page_wPK) <- rownames(camb_Adj) 
camb_page_wPK.sort <- sort(camb_page_wPK, T)
top_camb_page_wPK <- camb_page_wPK.sort[1:200]
```

#### Katz
```{r}
alpha <- 1/max(a$values) - (1/max(a$values))/100
camb_katz_wPK <- katz.centrality(camb_Adj, alpha, beta = IMP_deg)
names(camb_katz_wPK) <- rownames(camb_Adj) 
camb_katz_wPK.sort <- sort(camb_katz_wPK, T)
top_camb_katz_wPK <- camb_katz_wPK.sort[1:200]
```

#### EpistasisKatz
```{r}
beta <- diag(camb_reGAIN)
camb_regain.Katz <- camb_reGAIN
diag(camb_regain.Katz) <- 0
camb_ekatz_wPK <- katz.centrality(camb_regain.Katz, IMP_deg, beta)
names(camb_ekatz_wPK) <- rownames(camb_Adj) 
camb_ekatz_wPK.sort <- sort(camb_ekatz_wPK, T)
top_camb_EK_wPK <- camb_ekatz_wPK.sort[1:200]
```

#### EpistasisRank
```{r}
camb_ER_wPK <- EpistasisRank(camb_reGAIN, IMP_deg)
top_camb_ER_wPK <- camb_ER_wPK[1:200, ]
```

### Pathway Plots
```{r, fig.width=10, fig.height=8}
pathways <- read.csv("Reactome_pathways.csv", header = TRUE)

my_pathways_woPK <- cbind(names(top_camb_page_woPK), names(top_camb_katz_woPK), 
                          names(top_camb_EK_woPK), as.character(top_camb_ER_woPK$gene), colnames(top_camb_cov))
colnames(my_pathways_woPK) <- c("Page_woPK", "Katz_woPK", "EK_woPK", "ER_woPK", "CoV")
save(my_pathways_woPK, file = "my_pathways_woPK.RData")
pathway_mat_woPK <- matrix(0, 5, 9)
for (i in 1:5){
  for (j in 1:9){
    pathway_mat_woPK[i, j] <- length(intersect(my_pathways_woPK[, i], as.character(pathways[, j])))
  }
}

my_pathways_wPK <- cbind(names(top_camb_page_wPK), names(top_camb_katz_wPK), 
                         names(top_camb_EK_wPK), as.character(top_camb_ER_wPK$gene))

colnames(my_pathways_wPK) <- c("Page_wPK", "Katz_wPK", "EK_wPK", "ER_wPK")
save(my_pathways_wPK, file = "my_pathways_wPK.RData")
pathway_mat_wPK <- matrix(0, 5, 9)
for (i in 1:4){
  for (j in 1:9){
    pathway_mat_wPK[i, j] <- length(intersect(my_pathways_wPK[, i], as.character(pathways[, j])))
  }
}

plot_df <- data.frame(PK_Status=c("woPK", "woPK", "woPK", "woPK", "woPK", "wPK", "wPK", "wPK", "wPK", "wPK"), 
                      Method=c("PR", "Katz", "EK", "ER", "CoV", "PR", "Katz", "EK", "ER", "CoV"), 
                      rbind(pathway_mat_woPK, pathway_mat_wPK))

p1 <- ggplot(data=plot_df, aes(x=Method, y=X1, fill=PK_Status)) +
  labs(title="INHIBITION OF VOLTAGE GATED \n CD CA2 CHANNELS", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p2 <- ggplot(data=plot_df, aes(x=Method, y=X2, fill=PK_Status)) +
  labs(title="NEURONAL SYSTEM", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p3 <- ggplot(data=plot_df, aes(x=Method, y=X3, fill=PK_Status)) +
  labs(title="POTASSIUM CHANNELS", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p4 <- ggplot(data=plot_df, aes(x=Method, y=X4, fill=PK_Status)) +
  labs(title="TRANSMISSION ACROSS CHEMICAL \n SYNAPSES", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p5 <- ggplot(data=plot_df, aes(x=Method, y=X5, fill=PK_Status)) +
  labs(title="GPCR DOWNSTREAM SIGNALING", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p6 <- ggplot(data=plot_df, aes(x=Method, y=X6, fill=PK_Status)) +
  labs(title="SLC MEDICATED TRANSMEMBRANCE \n TRANSPORT", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p7 <- ggplot(data=plot_df, aes(x=Method, y=X7, fill=PK_Status)) +
  labs(title="NEUROTRANSMITTER RECEPTOR \n BINDING", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p8 <- ggplot(data=plot_df, aes(x=Method, y=X8, fill=PK_Status)) +
  labs(title="RAS ACTIVATION UPON CA2 INFLUX \n THROUGH NMDA RECEPTOR", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(axis.title = element_text(size = .5)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

p9 <- ggplot(data=plot_df, aes(x=Method, y=X9, fill=PK_Status)) +
  labs(title ="SIGNALLING BY GPCR", x="Ranking Method", y = "Number of Overlap Genes")+
  theme(plot.title = element_text(size = .2)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_brewer(palette="Paired") + theme_minimal() +
  scale_fill_manual(values=c("lightgray", "black")) +
  theme_classic()

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, nrow = 3, ncol = 3)
```

## Classification without prior knowledge
### PageRank
```{r}
# page_woPK_features <- NULL
# page_woPK_train_acc <- NULL
# page_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   page_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # Adjacency matrix to graph
#     camb_Adj_g <- graph.adjacency(camb_Adj_partial)
#     # page_woPKRank centrality
#     camb_page_woPK <- page.rank(camb_Adj_g)$vector
#     camb_page_woPK.sort <- sort(camb_page_woPK, T)
#     top_camb_page_woPK <- camb_page_woPK.sort[1:200]
#     page_woPK_inner.features <- c(page_woPK_inner.features, top_camb_page_woPK)
#     page_woPK_features <- c(page_woPK_features, top_camb_page_woPK)
# 
#   }
#   camb_page_woPK.expr <- camb_fltr.expr[, unique(as.character(names(page_woPK_inner.features)))] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_page_woPK.expr), label = camb_class)
#   camb_page_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
#                                         nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                         subsample = 0.5, objective = "binary:logistic")
#   
#   camb_page_woPK_bst.pred <- predict(camb_page_woPK_bst.model, as.matrix(camb_page_woPK.expr))
#   page_woPK_train_acc <- c(page_woPK_train_acc, accuracy(camb_class, camb_page_woPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_page_woPK.expr <- jap_fltr.expr[, unique(as.character(names(page_woPK_inner.features)))]
#   jap_page_woPK_bst.pred <- predict(camb_page_woPK_bst.model, as.matrix(jap_page_woPK.expr))
#   page_woPK_test_acc <- c(page_woPK_test_acc, accuracy(jap_class, jap_page_woPK_bst.pred)$prop.correct)
# }
load("page_woPK_train_acc.RData")
load("page_woPK_test_acc.RData")
pr_wo_train <- page_woPK_train_acc[which.min(page_woPK_train_acc - page_woPK_test_acc)]
pr_wo_test <- page_woPK_test_acc[which.min(page_woPK_train_acc - page_woPK_test_acc)]
```


### Katz
```{r}
# katz_woPK_features <- NULL
# katz_woPK_train_acc <- NULL
# katz_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   katz_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # katz_woPK centrality
#     a <- eigen(camb_Adj_partial)
#     alpha = signif(1/max(a$values), digits = 2)
#     beta <- rep(1, nrow(camb_Adj_partial))/nrow(camb_Adj_partial)
#     camb_katz_woPK <- katz.centrality(camb_Adj_partial, alpha, beta)
#     names(camb_katz_woPK) <- colnames(camb_Adj_partial)
#     camb_katz_woPK.sort <- sort(camb_katz_woPK, T)
#     top_camb_katz_woPK <- camb_katz_woPK.sort[1:200]
#     katz_woPK_inner.features <- c(katz_woPK_inner.features, top_camb_katz_woPK)
#     katz_woPK_features <- c(katz_woPK_features, top_camb_katz_woPK)
#   }
#   camb_katz_woPK.expr <- camb_fltr.expr[, unique(as.character(names(katz_woPK_inner.features)))] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_katz_woPK.expr), label = camb_class)
#   camb_katz_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
#                                         nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                         subsample = 0.5, objective = "binary:logistic")
#   
#   camb_katz_woPK_bst.pred <- predict(camb_katz_woPK_bst.model, as.matrix(camb_katz_woPK.expr))
#   katz_woPK_train_acc <- c(katz_woPK_train_acc, accuracy(camb_class, camb_katz_woPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_katz_woPK.expr <- jap_fltr.expr[, unique(as.character(names(katz_woPK_inner.features)))]
#   jap_katz_woPK_bst.pred <- predict(camb_katz_woPK_bst.model, as.matrix(jap_katz_woPK.expr))
#   katz_woPK_test_acc <- c(katz_woPK_test_acc, accuracy(jap_class, jap_katz_woPK_bst.pred)$prop.correct)
# }
load("katz_woPK_train_acc.RData")
load("katz_woPK_test_acc.RData")
k_wo_train <- katz_woPK_train_acc[which.min(katz_woPK_train_acc - katz_woPK_test_acc)]
k_wo_test <- katz_woPK_test_acc[which.min(katz_woPK_train_acc - katz_woPK_test_acc)]
```

### Epistasis-Katz
```{r}
# ekatz_woPK_features <- NULL
# ekatz_woPK_train_acc <- NULL
# ekatz_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ekatz_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]   
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     library(Rinbix)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- Rinbix::regainInbix(regain_matrix)
#     # ekatz_woPK centrality
#     alpha = 1/mean(colSums(camb_Regain$reGAIN))
#     beta = diag(camb_Regain$reGAIN)
#     diag(camb_Regain$reGAIN) <- 0
#     camb_ekatz_woPK <- katz.centrality(camb_Regain$reGAIN, alpha, beta)
#     names(camb_ekatz_woPK) <- colnames(camb_Regain$reGAIN)
#     camb_ekatz_woPK.sort <- sort(camb_ekatz_woPK, T)
#     top_camb_ekatz_woPK <- camb_ekatz_woPK.sort[1:200]
#     ekatz_woPK_inner.features <- c(ekatz_woPK_inner.features, top_camb_ekatz_woPK)
#     ekatz_woPK_features <- c(ekatz_woPK_features, top_camb_katz)
#   }
#   camb_ekatz_woPK.expr <- camb_fltr.expr[, unique(as.character(names(ekatz_woPK_inner.features)))] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ekatz_woPK.expr), label = camb_class)
#   camb_ekatz_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
#                                          nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                          subsample = 0.5, objective = "binary:logistic")
#   
#   camb_ekatz_woPK_bst.pred <- predict(camb_ekatz_woPK_bst.model, as.matrix(camb_ekatz_woPK.expr))
#   ekatz_woPK_train_acc <- c(ekatz_woPK_train_acc, accuracy(camb_class, camb_ekatz_woPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_ekatz_woPK.expr <- jap_fltr.expr[, unique(as.character(names(ekatz_woPK_inner.features)))]
#   jap_ekatz_woPK_bst.pred <- predict(camb_ekatz_woPK_bst.model, as.matrix(jap_ekatz_woPK.expr))
#   ekatz_woPK_test_acc <- c(ekatz_woPK_test_acc, accuracy(jap_class, jap_ekatz_woPK_bst.pred)$prop.correct)
# }
load("ekatz_woPK_train_acc.RData")
load("ekatz_woPK_test_acc.RData")
ek_wo_train <- ekatz_woPK_train_acc[which.min(ekatz_woPK_train_acc - ekatz_woPK_test_acc)]
ek_wo_test <- ekatz_woPK_test_acc[which.min(ekatz_woPK_train_acc - ekatz_woPK_test_acc)]
```

### EpistasisRank
```{r}
# ER_woPK_features <- NULL
# ER_woPK_train_acc <- NULL
# ER_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ER_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     library(Rinbix)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- regainInbix(regain_matrix)
#     # ER_woPKrank centrality
#     camb_ER_woPKrank <- snprankInbix(camb_Regain$reGAIN, gamma = .85)
#     top_camb_ER_woPK <- camb_ER_woPKrank[1:200, ]
#     ER_woPK_inner.features <- c(ER_woPK_inner.features, as.character(top_camb_ER_woPK$gene))
#     ER_woPK_features <- c(ER_woPK_features, as.character(top_camb_ER_woPK$gene))
#   }
#   camb_ER_woPK.expr <- camb_fltr.expr[, unique(ER_woPK_inner.features)] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ER_woPK.expr), label = camb_class)
#   camb_ER_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
#                                       nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                       subsample = 0.5, objective = "binary:logistic")
#   
#   camb_ER_woPK_bst.pred <- predict(camb_ER_woPK_bst.model, as.matrix(camb_ER_woPK.expr))
#   ER_woPK_train_acc <- c(ER_woPK_train_acc, accuracy(camb_class, camb_ER_woPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_ER_woPK.expr <- jap_fltr.expr[, unique(ER_woPK_inner.features)]
#   jap_ER_woPK_bst.pred <- predict(camb_ER_woPK_bst.model, as.matrix(jap_ER_woPK.expr))
#   ER_woPK_test_acc <- c(ER_woPK_test_acc, accuracy(jap_class, jap_ER_woPK_bst.pred)$prop.correct)
# }
load("ER_woPK_train_acc.RData")
load("ER_woPK_test_acc.RData")
er_wo_train <- ER_woPK_train_acc[which.min(ER_woPK_train_acc - ER_woPK_test_acc)]
er_wo_test <- ER_woPK_test_acc[which.min(ER_woPK_train_acc - ER_woPK_test_acc)]
```

## Classification with prior knowledge
### PageRank
```{r}
# page_wPK_features <- NULL
# page_wPK_train_acc <- NULL
# page_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   page_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # Adjacency matrix to graph
#     camb_Adj_g <- graph.adjacency(camb_Adj_partial)
#     # page_wPKRank centrality
#     camb_page_wPK <- as.numeric(my.pagerank(camb_Adj_partial, IMP_deg))
#     names(camb_page_wPK) <- rownames(camb_Adj_partial) 
#     camb_page_wPK.sort <- sort(camb_page_wPK, T)
#     top_camb_page_wPK <- camb_page_wPK.sort[1:200]
#     page_wPK_inner.features <- c(page_wPK_inner.features, as.character(names(top_camb_page_wPK)))
#     page_wPK_features <- c(page_wPK_features, as.character(names(top_camb_page_wPK)))
#   }
#   camb_page_wPK.expr <- camb_fltr.expr[, unique(page_wPK_inner.features)] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_page_wPK.expr), label = camb_class)
#   camb_page_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
#                                        nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                        subsample = 0.5, objective = "binary:logistic")
#   
#   camb_page_wPK_bst.pred <- predict(camb_page_wPK_bst.model, as.matrix(camb_page_wPK.expr))
#   page_wPK_train_acc <- c(page_wPK_train_acc, accuracy(camb_class, camb_page_wPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_page_wPK.expr <- jap_fltr.expr[, unique(page_wPK_inner.features)]
#   jap_page_wPK_bst.pred <- predict(camb_page_wPK_bst.model, as.matrix(jap_page_wPK.expr))
#   page_wPK_test_acc <- c(page_wPK_test_acc, accuracy(jap_class, jap_page_wPK_bst.pred)$prop.correct)
# }
load("page_wPK_train_acc.RData")
load("page_wPK_test_acc.RData")
pr_w_train <- page_wPK_train_acc[which.min(page_wPK_train_acc - page_wPK_test_acc)]
pr_w_test <- page_wPK_test_acc[which.min(page_wPK_train_acc - page_wPK_test_acc)]
```

### Katz
```{r}
# katz_wPK_features <- NULL
# katz_wPK_train_acc <- NULL
# katz_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   katz_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # katz_wPK centrality
#     a <- eigen(camb_Adj_partial)
#     alpha <- signif(1/max(a$values), digits = 2)
#     camb_katz_wPK <- katz.centrality(camb_Adj_partial, alpha, beta = IMP_deg)
#     names(camb_katz_wPK) <- rownames(camb_Adj_partial) 
#     camb_katz_wPK.sort <- sort(camb_katz_wPK, T)
#     top_camb_katz_wPK <- camb_katz_wPK.sort[1:200]
#     katz_wPK_inner.features <- c(katz_wPK_inner.features, as.character(names(top_camb_katz_wPK)))
#     katz_wPK_features <- c(katz_wPK_features, as.character(names(top_camb_katz_wPK)))
#   }
#   camb_katz_wPK.expr <- camb_fltr.expr[, unique(katz_wPK_inner.features)] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_katz_wPK.expr), label = camb_class)
#   camb_katz_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
#                                        nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                        subsample = 0.7, objective = "binary:logistic")
#   
#   camb_katz_wPK_bst.pred <- predict(camb_katz_wPK_bst.model, as.matrix(camb_katz_wPK.expr))
#   katz_wPK_train_acc <- c(katz_wPK_train_acc, accuracy(camb_class, camb_katz_wPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_katz_wPK.expr <- jap_fltr.expr[, unique(katz_wPK_inner.features)]
#   jap_katz_wPK_bst.pred <- predict(camb_katz_wPK_bst.model, as.matrix(jap_katz_wPK.expr))
#   katz_wPK_test_acc <- c(katz_wPK_test_acc, accuracy(jap_class, jap_katz_wPK_bst.pred)$prop.correct)
# }
load("katz_wPK_train_acc.RData")
load("katz_wPK_test_acc.RData")
k_w_train <- katz_wPK_train_acc[which.min(katz_wPK_train_acc - katz_wPK_test_acc)]
k_w_test <- katz_wPK_test_acc[which.min(katz_wPK_train_acc - katz_wPK_test_acc)]
```

### Epistasis-Katz
```{r}
# ekatz_wPK_features <- NULL
# ekatz_wPK_train_acc <- NULL
# ekatz_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ekatz_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     library(Rinbix)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- regainInbix(regain_matrix)
#     # ekatz_wPK centrality
#     beta <- diag(camb_Regain$reGAIN)
#     diag(camb_Regain$reGAIN) <- 0
#     camb_ekatz_wPK <- katz.centrality(camb_Regain$reGAIN, IMP_deg, beta)
#     names(camb_ekatz_wPK) <- colnames(camb_expr) 
#     camb_ekatz_wPK.sort <- sort(camb_ekatz_wPK, T)
#     top_camb_ekatz_wPK <- camb_ekatz_wPK.sort[1:200]
#     ekatz_wPK_inner.features <- c(ekatz_wPK_inner.features, as.character(names(top_camb_ekatz_wPK)))
#     ekatz_wPK_features <- c(ekatz_wPK_features, as.character(names(top_camb_ekatz_wPK)))
#   }
#   camb_ekatz_wPK.expr <- camb_fltr.expr[, unique(ekatz_wPK_inner.features)] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ekatz_wPK.expr), label = camb_class)
#   camb_ekatz_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=2, eta = .1, nthread = 2, 
#                                         nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                         subsample = 0.4, objective = "binary:logistic")
#   
#   camb_ekatz_wPK_bst.pred <- predict(camb_ekatz_wPK_bst.model, as.matrix(camb_ekatz_wPK.expr))
#   ekatz_wPK_train_acc <- c(ekatz_wPK_train_acc, accuracy(camb_class, camb_ekatz_wPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_ekatz_wPK.expr <- jap_fltr.expr[, unique(ekatz_wPK_inner.features)]
#   jap_ekatz_wPK_bst.pred <- predict(camb_ekatz_wPK_bst.model, as.matrix(jap_ekatz_wPK.expr))
#   ekatz_wPK_test_acc <- c(ekatz_wPK_test_acc, accuracy(jap_class, jap_ekatz_wPK_bst.pred)$prop.correct)
# }
load("ekatz_wPK_train_acc.RData")
load("ekatz_wPK_test_acc.RData")
ek_w_train <- ekatz_wPK_train_acc[which.min(ekatz_wPK_train_acc - ekatz_wPK_test_acc)]
ek_w_test <- ekatz_wPK_test_acc[which.min(ekatz_wPK_train_acc - ekatz_wPK_test_acc)]
```

### EpistasisRank
```{r}
# ER_wPK_features <- NULL
# ER_wPK_train_acc <- NULL
# ER_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ER_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     library(Rinbix)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- regainInbix(regain_matrix)
#     # ER_wPKrank centrality
#     camb_ER_wPKrank <- EpistasisRank(camb_Regain$reGAIN, IMP_deg)
#     top_camb_ER_wPK <- camb_ER_wPKrank[1:200,]
#     ER_wPK_inner.features <- c(ER_wPK_inner.features, as.character(top_camb_ER_wPK$gene))
#     ER_wPK_features <- c(ER_wPK_features, as.character(top_camb_ER_wPK$gene))
#   }
#   camb_ER_wPK.expr <- camb_fltr.expr[, unique(ER_wPK_inner.features)] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ER_wPK.expr), label = camb_class)
#   camb_ER_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=2, eta = .1, nthread = 2, 
#                                      nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                      subsample = 0.7, objective = "binary:logistic")
#   
#   camb_ER_wPK_bst.pred <- predict(camb_ER_wPK_bst.model, as.matrix(camb_ER_wPK.expr))
#   ER_wPK_train_acc <- c(ER_wPK_train_acc, accuracy(camb_class, camb_ER_wPK_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_ER_wPK.expr <- jap_fltr.expr[, unique(ER_wPK_inner.features)]
#   jap_ER_wPK_bst.pred <- predict(camb_ER_wPK_bst.model, as.matrix(jap_ER_wPK.expr))
#   ER_wPK_test_acc <- c(ER_wPK_test_acc, accuracy(jap_class, jap_ER_wPK_bst.pred)$prop.correct)
# }
load("ER_wPK_train_acc.RData")
load("ER_wPK_test_acc.RData")
er_w_train <- ER_wPK_train_acc[which.min(ER_wPK_train_acc - ER_wPK_test_acc)]
er_w_test <- ER_wPK_test_acc[which.min(ER_wPK_train_acc - ER_wPK_test_acc)]
```

### CoV filtering
```{r}
# cov_features <- NULL
# cov_train_acc <- NULL
# cov_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   cov_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]] 
#     # CV filtering
#     exprData <- t(camb_fltr.expr[fold_idx, ])
#     thresh <- .5
#     camb_expr.fltr_cov <- camb_fltr.expr
#     while (ncol(camb_expr.fltr_cov)>200){
#       camb_expr.fltr_cov <- t(cov.filter(exprData,threshold = thresh)$fdata)
#       thresh <- thresh-0.001
#     }
#     cov_inner.features <- c(cov_inner.features, colnames(camb_expr.fltr_cov))
#     cov_features <- c(cov_features, colnames(camb_expr.fltr_cov))
#   }
#   camb_cov.expr <- camb_fltr.expr[, unique(as.character(cov_inner.features))] 
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_cov.expr), label = camb_class)
#   camb_cov_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
#                                   nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
#                                   subsample = 0.5, objective = "binary:logistic")
#   
#   camb_cov_bst.pred <- predict(camb_cov_bst.model, as.matrix(camb_cov.expr))
#   cov_train_acc <- c(cov_train_acc, accuracy(camb_class, camb_cov_bst.pred)$prop.correct)
#   
#   # Testing
#   jap_cov.expr <- jap_fltr.expr[, unique(as.character(cov_inner.features))]
#   jap_cov_bst.pred <- predict(camb_cov_bst.model, as.matrix(jap_cov.expr))
#   cov_test_acc <- c(cov_test_acc, accuracy(jap_class, jap_cov_bst.pred)$prop.correct)
# }
load("cov_train_acc.RData")
load("cov_test_acc.RData")
cov_train <- cov_train_acc[which.min(cov_train_acc - cov_test_acc)]
cov_test <- cov_test_acc[which.min(cov_train_acc - cov_test_acc)]
```

## XGBoost
```{r}
# Training
camb_dtrain <- xgb.DMatrix(as.matrix(camb_fltr.expr), label = camb_class)
camb_xgb_model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2, 
                                   nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
                                   subsample = 0.5, objective = "binary:logistic")

camb_xgb_pred <- predict(camb_xgb_model, as.matrix(camb_fltr.expr))
xgb_train <- accuracy(camb_class, camb_xgb_pred)$prop.correct

# Testing
jap_xgb_pred <- predict(camb_xgb_model, as.matrix(jap_fltr.expr))
xgb_test <- accuracy(jap_class, jap_xgb_pred)$prop.correct
```


## Accuracy Plots
```{r, fig.width=6, fig.height=6}
PR.Katz_df <- data.frame(
  Type = factor(c("woPK", "wPK", "woPK", "wPK", "woPK", "wPK", "woPK", "wPK"), levels = c("woPK", "wPK")),
  Method = factor(c("PageRank", "PageRank", "Katz", "Katz", "PageRank", "PageRank", "Katz", "Katz"), 
                  levels = c("PageRank", "Katz")),
  Accuracy = c(pr_wo_train, pr_w_train, k_wo_train, k_w_train, pr_wo_test, pr_w_test, k_wo_test, k_w_test),
  Status  = c("Cambridge-Train", "Cambridge-Train", "Cambridge-Train", "Cambridge-Train", 
              "Japan-Test", "Japan-Test", "Japan-Test", "Japan-Test")
)

EK.ER_df <- data.frame(
  Type = factor(c("woPK", "wPK", "woPK", "wPK", "woPK", "wPK", "woPK", "wPK"), levels = c("woPK", "wPK")),
  Method = factor(c("Epistasis-Katz", "Epistasis-Katz", "EpistasisRank", "EpistasisRank", 
                    "Epistasis-Katz", "Epistasis-Katz", "EpistasisRank", "EpistasisRank"), 
                  levels = c("Epistasis-Katz", "EpistasisRank")),
  Accuracy = c(ek_wo_train, ek_w_train, er_wo_train, er_w_train, ek_wo_test, ek_w_test, er_wo_test, er_w_test),
  Status = c("Cambridge-Train", "Cambridge-Train", "Cambridge-Train", "Cambridge-Train",
             "Japan-Test", "Japan-Test", "Japan-Test", "Japan-Test")
)

cov_pec_df <- data.frame(
  Type = factor(c("woPK", "woPK", "woPK", "woPK")),
  Method = factor(c("CoV", "xgb", "CoV", "xgb"), levels = c("CoV", "xgb")),
  Accuracy = c(cov_train, xgb_train, cov_test, xgb_test),
  Status = c("Cambridge-Train", "Cambridge-Train", "Japan-Test", "Japan-Test")
)


p1 <- ggplot(PR.Katz_df, aes(x = Type, y = Accuracy, group = Method)) +
  geom_point() + geom_line(aes(linetype=Method)) +
  geom_point(data = cov_pec_df, mapping = aes(x = Type, y = Accuracy, shape = Method)) +
  scale_linetype_manual(values=c("twodash", "solid"))+
  facet_wrap(~Status) +
  theme(legend.position="top") +
  guides(group = guide_legend(order=1),
         shape = guide_legend(order=2))
p2 <- ggplot(EK.ER_df, aes(x = Type, y = Accuracy, group = Method)) +
  geom_point() + geom_line(aes(linetype=Method)) +
  geom_point(data = cov_pec_df, mapping = aes(x = Type, y = Accuracy, shape = Method)) +
  scale_linetype_manual(values=c("twodash", "solid"))+
  facet_wrap(~Status) +
  theme(legend.position="top") +
  guides(group = guide_legend(order=1),
         shape = guide_legend(order=2))
grid.arrange(p1, p2, nrow = 2)
```
