---
title: "Classification (with and without prior knowledge) Example"
author: "Saeid Parvandeh"
date: "10/14/2018"
output:
  github_document:
  keep_md: yes
knitr:
  clean: no
run_pandoc: no
---

```{r setup, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

### Install required packages
```{r results='hide', message=FALSE, warning=FALSE}
check.packages <- function(pkg){
  # check.packages function: install and load multiple R packages.
  # Check to see if packages are installed. Install them if they are not, 
  # then load them into the R session.
  # https://gist.github.com/smithdanielle/9913897
  
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, repos = "http://cran.us.r-project.org", dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}


if (!("privateEC" %in% installed.packages()[,"Package"])){
  devtools::install_github("insilico/privateEC", build_vignettes = TRUE)
}
library(privateEC)

if (!("devtools" %in% installed.packages()[,"Package"])){
  install.packages("devtools", repos = "http://cran.us.r-project.org", dependencies = TRUE)
}
library(devtools)

# install github package
if (!("Rinbix" %in% installed.packages()[,"Package"])){
  devtools::install_github("insilico/Rinbix", build_vignettes = TRUE)
}

# install cran packages
packages <- c("ggplot2", "CORElearn", "gridExtra", "lattice", "caret", "glmnet", "xgboost", "igraph",
              "gdata", "SDMTools", "RMariaDB")

# # install bioconductor packages
# source("https://bioconductor.org/biocLite.R")
# biocLite("GEOquery")
# biocLite("Biobase")
# biocLite("preprocessCore")

source("impquery-functions.R")
```

### Loading data from GEO website
```{r}
# # get data from GEO
# Elist <- getGEO("GSE98793")
# 
# # Elicit gene expression
# expr <- exprs(Elist[[1]])
# # Elicit the titles (MDD and CON)
# titles <- pData(Elist[[1]])[, 1]
# # assign the status labels
# camb_class <- ifelse(grepl("control", titles), 0, 1)
# # Making sure that probIDs and rownames are identical
# probID <- Elist$GSE98793_series_matrix.txt.gz@featureData@data$ID
# identical(as.character(probID), rownames(expr))
# # Substitute probID with gene symboles
# gene.symbol <- Elist$GSE98793_series_matrix.txt.gz@featureData@data$`Gene Symbol`
# rownames(expr) <- gene.symbol
# expr <- expr[-which(rownames(expr)==""), ]
# 
# # z-transformation on control samples
# ctrl_idx <- which(camb_class==0)
# ctrl_expr <- expr[, ctrl_idx]
# z.trans_mat <- matrix(0,nrow = nrow(expr),ncol=ncol(expr))
# for (i in 1:nrow(ctrl_expr)){
#   z.trans_mat[i, ] <- pnorm(as.numeric(expr[i, ]),
#                             mean(as.numeric(ctrl_expr[i, ])),
#                             sd(as.numeric(ctrl_expr[i, ])))
# }
# 
# camb_expr <- t(z.trans_mat)
# colnames(camb_expr) <- rownames(expr)
# rownames(camb_expr) <- colnames(expr)
# 
# save(camb_expr, file = "camb_expr.RData")
# save(camb_class, file = "camb_class.RData")

# # get data from GEO
# Elist <- getGEO("GSE76826")
# 
# # Elicit gene expression
# expr <- exprs(Elist[[1]])
# # Elicit the titles (MDD and HC)
# titles <- pData(Elist[[1]])[, 1]
# # assign the status labels
# jap_class <- ifelse(grepl("HC", titles), 0, 1)
# # Making sure that probIDs and rownames are identical
# probID <- Elist$GSE76826_series_matrix.txt.gz@featureData@data$ID
# identical(as.character(probID), rownames(expr))
# # Substitute probID with gene symboles
# gene.symbol <- Elist$GSE76826_series_matrix.txt.gz@featureData@data$GENE_SYMBOL
# rownames(expr) <- gene.symbol
# expr <- expr[-which(rownames(expr)==""), ]
# 
# # z-transform in control samples
# ctrl_idx <- which(jap_class==0)
# ctrl_expr <- expr[, ctrl_idx]
# z.trans_mat <- matrix(0,nrow = nrow(expr),ncol=ncol(expr))
# for (i in 1:nrow(ctrl_expr)){
#   z.trans_mat[i, ] <- pnorm(as.numeric(expr[i, ]),
#                             mean(as.numeric(ctrl_expr[i, ])),
#                             sd(as.numeric(ctrl_expr[i, ])))
# }
# 
# jap_expr <- t(z.trans_mat)
# colnames(jap_expr) <- rownames(expr)
# rownames(jap_expr) <- colnames(expr)
# 
# save(jap_expr, file = "jap_expr.RData")
# save(jap_class, file = "jap_class.RData")
load("jap_class.RData")
```


### Subsampling
```{r}
# # balance the number of case/control subjects.
# case_idx <- which(camb_class==1)
# set.seed(123)
# half_case_idx <- sample(case_idx, length(case_idx)/2)
# camb_class <- camb_class[-half_case_idx]
# camb_expr <- camb_expr[-half_case_idx, ]
# save(camb_expr, file = "camb_expr_equalized.RData")
# save(camb_class, file = "camb_class_equalized.RData")

# # subsample the Cambridge data to the size of Japan data
# # case
# case_idx <- which(camb_class==1)
# set.seed(456)
# subsample_case_idx <- sample(case_idx, length(case_idx)-44)
# # control
# control_idx <- which(camb_class==0)
# subsample_control_idx <- sample(control_idx, length(control_idx)-44)
# camb_class <- camb_class[c(subsample_control_idx, subsample_case_idx)]
# save(camb_class, file = "camb_sub_class.RData")
load("camb_sub_class.RData")
# camb_expr <- camb_expr[c(subsample_control_idx,subsample_case_idx), ]
# save(camb_expr, file = "camb_sub_expr.RData")
load("camb_sub_expr.RData")
```


### Filter noisy genes using coefficient of variation
```{r}
# # Coefficient of Variation (CoV) filtering -- Cambridge, Japan
# exprData1 <- t(camb_expr)
# camb_expr.fltr_cov <- data.frame(t(Rinbix::geneLowCoefOfVarFilter(exprData1,.559095)$fdata)) 
# dim(camb_expr.fltr_cov)
# 
# exprData2 <- t(jap_expr)
# jap_expr.fltr_cov <- data.frame(t(Rinbix::geneLowCoefOfVarFilter(exprData2,.559095)$fdata)) 
# dim(jap_expr.fltr_cov)
# 
# camb_genes <- as.character(colnames(camb_expr.fltr_cov))
# jap_genes <- as.character(colnames(jap_expr.fltr_cov))
# 
# common_genes <- intersect(camb_genes, jap_genes)
# length(common_genes)
# camb_fltr.expr <- camb_expr.fltr_cov[, which(camb_genes %in% common_genes)]
# camb_fltr.expr <- camb_fltr.expr[, order(colnames(camb_fltr.expr))]
# dim(camb_fltr.expr)
# save(camb_fltr.expr, file = "camb_fltr.expr.RData")
load("camb_fltr.expr.RData")
# jap_fltr.expr <- jap_expr.fltr_cov[, which(jap_genes %in% common_genes)]
# jap_fltr.expr <- jap_fltr.expr[, order(colnames(jap_fltr.expr))]
# dim(jap_fltr.expr)
# save(jap_fltr.expr, file = "jap_fltr.expr.RData")
load("jap_fltr.expr.RData")
```

### Computing the IMP query network
```{r}
# # wirte all genes in a file
# write.matrix(colnames(camb_fltr.expr), file = "top5000uniq-genes.txt", sep = "\t")
# geneListFile <- "top5000uniq-genes.txt"
# outPrefix <-    "top5000uniq-genes"
# 
# cat("Gene list file:      ", geneListFile, "\n", sep = "\t")
# cat("Confidence threshold:", confidenceThreshold, "\n", sep = "\t")
# cat("Target network size: ", targetListSize, "\n", sep = "\t")
# cat("Output prefix:       ", outPrefix, "\n", sep = "\t")
# 
# # read the list of input genes
# cat("Reading gene list (network node) file\n")
# geneListTable <- read.table(geneListFile, header = FALSE, sep = "\t",
#                             stringsAsFactors = FALSE)
# numInputGenes <- nrow(geneListTable)
# inputGeneList <- as.character(geneListTable[, 1])
# # TODO: lists with main effects would have have scores here in column 2
# inputScoresList <- rep.int(x = 0, times = numInputGenes)
# cat(length(inputGeneList), " genes read \n")
# 
# # connect to the MySQL database
# cat("Opening a connection to the database\n")
# cursorImpDatabase <- RMariaDB::dbConnect(MariaDB(),
#                                          user = "anonymous",
#                                          password = "s3cr3t",
#                                          dbname = "imp",
#                                          host = "127.0.0.1")
# 
# # create the interactions matrix, optionally expanding the list first with IMP interactions
# cat("Building the matrix from database queries\n")
# hasMainEffects <- FALSE
# isRankedList <- FALSE
# outSifFilename <- paste(outPrefix, ".sif", sep = "")
# VERBOSE_FLAG <- FALSE
# confidenceThreshold <- 0  # confidence range 0.0 to 1.0
# targetListSize = 0 # use only original list
# matrixGeneList = fillRegain(cursorImp = cursorImpDatabase,
#                             geneList = inputGeneList,
#                             scoreList = inputScoresList,
#                             hasMainEffectScores = hasMainEffects,
#                             ranked = isRankedList,
#                             confidenceThreshold = confidenceThreshold,
#                             addMode = FALSE,
#                             sifFilename = outSifFilename,
#                             verbose = VERBOSE_FLAG)
# 
# # write output files
# # matrix
# outFilename <- paste(outPrefix, ".adjmat", sep = "")
# cat("Saving", outFilename, "\n")
# colnames(matrixGeneList$filledMatrix) <- matrixGeneList$filledGeneList
# write.table(matrixGeneList$filledMatrix, file = outFilename, sep = "\t",
#             col.names = TRUE, row.names = FALSE, quote = FALSE)
# 
# impNetwork <- as.matrix(read.table("top5000uniq-genes.adjmat", header = TRUE, sep = "\t"))
# # replace all real values with 1, so that we will have a binary matrix
# impNetwork[impNetwork > 0] <- 1
# impNetwork <- matrix(as.integer(impNetwork), nrow = 5000)
# rownames(impNetwork) <- colnames(camb_fltr.expr)
# colnames(impNetwork) <- colnames(camb_fltr.expr)
# save(impNetwork, file = "impNetwork.RData")

load("impNetwork.RData")
```

#### Create prior knowledge network
```{r}
# prior knowledge
impgraph <- igraph::graph_from_adjacency_matrix(impNetwork, mode = "undirected", diag = TRUE, add.rownames = NA)
IMP_deg <- igraph::degree(impgraph, normalized = TRUE, loops = FALSE)
```

### Computing the reGAIN matrix
```{r}
# # Cambridge
# camb_data <- data.frame(camb_fltr.expr, class = camb_class)
# camb_reGAIN <- Rinbix::regainInbix(camb_data)$reGAIN
# save(camb_reGAIN, file = "camb_reGAIN.RData")
load("camb_reGAIN.RData")
# # Japan
# jap_data <- data.frame(jap_fltr.expr, class = jap_class)
# jap_reGAIN <- Rinbix::regainInbix(jap_data)$reGAIN
# save(jap_reGAIN, file = "jap_reGAIN.RData")
load("jap_reGAIN.RData")
```


## Classification without prior knowledge
### PageRank
```{r}
# page_woPK_features <- NULL
# page_woPK_train_acc <- NULL
# page_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   page_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # Adjacency matrix to graph
#     camb_Adj_g <- graph.adjacency(camb_Adj_partial)
#     # page_woPKRank centrality
#     camb_page_woPK <- Rinbix::PageRank(camb_Adj_g)$vector
#     camb_page_woPK.sort <- sort(camb_page_woPK, T)
#     top_camb_page_woPK <- camb_page_woPK.sort[1:200]
#     page_woPK_inner.features <- c(page_woPK_inner.features, top_camb_page_woPK)
#     page_woPK_features <- c(page_woPK_features, top_camb_page_woPK)
# 
#   }
#   camb_page_woPK.expr <- camb_fltr.expr[, unique(as.character(names(page_woPK_inner.features)))]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_page_woPK.expr), label = camb_class)
#   camb_page_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2,
#                                         nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                         subsample = 0.5, objective = "binary:logistic")
# 
#   camb_page_woPK_bst.pred <- predict(camb_page_woPK_bst.model, as.matrix(camb_page_woPK.expr))
#   page_woPK_train_acc <- c(page_woPK_train_acc, accuracy(camb_class, camb_page_woPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_page_woPK.expr <- jap_fltr.expr[, unique(as.character(names(page_woPK_inner.features)))]
#   jap_page_woPK_bst.pred <- predict(camb_page_woPK_bst.model, as.matrix(jap_page_woPK.expr))
#   page_woPK_test_acc <- c(page_woPK_test_acc, accuracy(jap_class, jap_page_woPK_bst.pred)$prop.correct)
# }
load("page_woPK_train_acc.RData")
load("page_woPK_test_acc.RData")

pr_wo_train <- page_woPK_train_acc[which.min(page_woPK_train_acc - page_woPK_test_acc)]
pr_wo_test <- page_woPK_test_acc[which.min(page_woPK_train_acc - page_woPK_test_acc)]
```


### Katz
```{r}
# katz_woPK_features <- NULL
# katz_woPK_train_acc <- NULL
# katz_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   katz_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # katz_woPK centrality
#     a <- eigen(camb_Adj_partial)
#     alpha = signif(1/max(a$values), digits = 2)
#     beta <- rep(1, nrow(camb_Adj_partial))/nrow(camb_Adj_partial)
#     camb_katz_woPK <- Rinbix::EpistasisKatz(camb_Adj_partial, alpha, beta)
#     names(camb_katz_woPK) <- colnames(camb_Adj_partial)
#     camb_katz_woPK.sort <- sort(camb_katz_woPK, T)
#     top_camb_katz_woPK <- camb_katz_woPK.sort[1:200]
#     katz_woPK_inner.features <- c(katz_woPK_inner.features, top_camb_katz_woPK)
#     katz_woPK_features <- c(katz_woPK_features, top_camb_katz_woPK)
#   }
#   camb_katz_woPK.expr <- camb_fltr.expr[, unique(as.character(names(katz_woPK_inner.features)))]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_katz_woPK.expr), label = camb_class)
#   camb_katz_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2,
#                                         nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                         subsample = 0.5, objective = "binary:logistic")
# 
#   camb_katz_woPK_bst.pred <- predict(camb_katz_woPK_bst.model, as.matrix(camb_katz_woPK.expr))
#   katz_woPK_train_acc <- c(katz_woPK_train_acc, accuracy(camb_class, camb_katz_woPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_katz_woPK.expr <- jap_fltr.expr[, unique(as.character(names(katz_woPK_inner.features)))]
#   jap_katz_woPK_bst.pred <- predict(camb_katz_woPK_bst.model, as.matrix(jap_katz_woPK.expr))
#   katz_woPK_test_acc <- c(katz_woPK_test_acc, accuracy(jap_class, jap_katz_woPK_bst.pred)$prop.correct)
# }
load("katz_woPK_train_acc.RData")
load("katz_woPK_test_acc.RData")

k_wo_train <- katz_woPK_train_acc[which.min(katz_woPK_train_acc - katz_woPK_test_acc)]
k_wo_test <- katz_woPK_test_acc[which.min(katz_woPK_train_acc - katz_woPK_test_acc)]
```

### Epistasis-Katz
```{r}
# ekatz_woPK_features <- NULL
# ekatz_woPK_train_acc <- NULL
# ekatz_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ekatz_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- Rinbix::regainInbix(regain_matrix)
#     # ekatz_woPK centrality
#     alpha = 1/mean(colSums(camb_Regain$reGAIN))
#     beta = diag(camb_Regain$reGAIN)
#     diag(camb_Regain$reGAIN) <- 0
#     camb_ekatz_woPK <- Rinbix::EpistasisKatz(camb_Regain$reGAIN, alpha, beta)
#     names(camb_ekatz_woPK) <- colnames(camb_Regain$reGAIN)
#     camb_ekatz_woPK.sort <- sort(camb_ekatz_woPK, T)
#     top_camb_ekatz_woPK <- camb_ekatz_woPK.sort[1:200]
#     ekatz_woPK_inner.features <- c(ekatz_woPK_inner.features, top_camb_ekatz_woPK)
#     ekatz_woPK_features <- c(ekatz_woPK_features, top_camb_katz)
#   }
#   camb_ekatz_woPK.expr <- camb_fltr.expr[, unique(as.character(names(ekatz_woPK_inner.features)))]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ekatz_woPK.expr), label = camb_class)
#   camb_ekatz_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2,
#                                          nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                          subsample = 0.5, objective = "binary:logistic")
# 
#   camb_ekatz_woPK_bst.pred <- predict(camb_ekatz_woPK_bst.model, as.matrix(camb_ekatz_woPK.expr))
#   ekatz_woPK_train_acc <- c(ekatz_woPK_train_acc, accuracy(camb_class, camb_ekatz_woPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_ekatz_woPK.expr <- jap_fltr.expr[, unique(as.character(names(ekatz_woPK_inner.features)))]
#   jap_ekatz_woPK_bst.pred <- predict(camb_ekatz_woPK_bst.model, as.matrix(jap_ekatz_woPK.expr))
#   ekatz_woPK_test_acc <- c(ekatz_woPK_test_acc, accuracy(jap_class, jap_ekatz_woPK_bst.pred)$prop.correct)
# }
load("ekatz_woPK_train_acc.RData")
load("ekatz_woPK_test_acc.RData")

ek_wo_train <- ekatz_woPK_train_acc[which.min(ekatz_woPK_train_acc - ekatz_woPK_test_acc)]
ek_wo_test <- ekatz_woPK_test_acc[which.min(ekatz_woPK_train_acc - ekatz_woPK_test_acc)]
```

### EpistasisRank
```{r}
# ER_woPK_features <- NULL
# ER_woPK_train_acc <- NULL
# ER_woPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ER_woPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- Rinbix::regainInbix(regain_matrix)
#     # ER_woPKrank centrality
#     camb_ER_woPKrank <- snprankInbix(camb_Regain$reGAIN, gamma = .85)
#     top_camb_ER_woPK <- camb_ER_woPKrank[1:200, ]
#     ER_woPK_inner.features <- c(ER_woPK_inner.features, as.character(top_camb_ER_woPK$gene))
#     ER_woPK_features <- c(ER_woPK_features, as.character(top_camb_ER_woPK$gene))
#   }
#   camb_ER_woPK.expr <- camb_fltr.expr[, unique(ER_woPK_inner.features)]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ER_woPK.expr), label = camb_class)
#   camb_ER_woPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2,
#                                       nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                       subsample = 0.5, objective = "binary:logistic")
# 
#   camb_ER_woPK_bst.pred <- predict(camb_ER_woPK_bst.model, as.matrix(camb_ER_woPK.expr))
#   ER_woPK_train_acc <- c(ER_woPK_train_acc, accuracy(camb_class, camb_ER_woPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_ER_woPK.expr <- jap_fltr.expr[, unique(ER_woPK_inner.features)]
#   jap_ER_woPK_bst.pred <- predict(camb_ER_woPK_bst.model, as.matrix(jap_ER_woPK.expr))
#   ER_woPK_test_acc <- c(ER_woPK_test_acc, accuracy(jap_class, jap_ER_woPK_bst.pred)$prop.correct)
# }
load("ER_woPK_train_acc.RData")
load("ER_woPK_test_acc.RData")

er_wo_train <- ER_woPK_train_acc[which.min(ER_woPK_train_acc - ER_woPK_test_acc)]
er_wo_test <- ER_woPK_test_acc[which.min(ER_woPK_train_acc - ER_woPK_test_acc)]
```

## Classification with prior knowledge
### PageRank
```{r}
# page_wPK_features <- NULL
# page_wPK_train_acc <- NULL
# page_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   page_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # Adjacency matrix to graph
#     camb_Adj_g <- graph.adjacency(camb_Adj_partial)
#     # page_wPKRank centrality
#     camb_page_wPK <- as.numeric(Rinbix::PageRank(camb_Adj_partial, IMP_deg))
#     names(camb_page_wPK) <- rownames(camb_Adj_partial)
#     camb_page_wPK.sort <- sort(camb_page_wPK, T)
#     top_camb_page_wPK <- camb_page_wPK.sort[1:200]
#     page_wPK_inner.features <- c(page_wPK_inner.features, as.character(names(top_camb_page_wPK)))
#     page_wPK_features <- c(page_wPK_features, as.character(names(top_camb_page_wPK)))
#   }
#   camb_page_wPK.expr <- camb_fltr.expr[, unique(page_wPK_inner.features)]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_page_wPK.expr), label = camb_class)
#   camb_page_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2,
#                                        nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                        subsample = 0.5, objective = "binary:logistic")
# 
#   camb_page_wPK_bst.pred <- predict(camb_page_wPK_bst.model, as.matrix(camb_page_wPK.expr))
#   page_wPK_train_acc <- c(page_wPK_train_acc, accuracy(camb_class, camb_page_wPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_page_wPK.expr <- jap_fltr.expr[, unique(page_wPK_inner.features)]
#   jap_page_wPK_bst.pred <- predict(camb_page_wPK_bst.model, as.matrix(jap_page_wPK.expr))
#   page_wPK_test_acc <- c(page_wPK_test_acc, accuracy(jap_class, jap_page_wPK_bst.pred)$prop.correct)
# }
load("page_wPK_train_acc.RData")
load("page_wPK_test_acc.RData")

pr_w_train <- page_wPK_train_acc[which.min(page_wPK_train_acc - page_wPK_test_acc)]
pr_w_test <- page_wPK_test_acc[which.min(page_wPK_train_acc - page_wPK_test_acc)]
```

### Katz
```{r}
# katz_wPK_features <- NULL
# katz_wPK_train_acc <- NULL
# katz_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   katz_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # correlation matrix
#     camb_expre.cor <- cor(camb_fltr.expr[fold_idx,])
#     # co-expression to binary
#     camb_Adj_partial <- ifelse(abs(camb_expre.cor) > .15, 1, 0)
#     diag(camb_Adj_partial) <- 0
#     # katz_wPK centrality
#     a <- eigen(camb_Adj_partial)
#     alpha <- signif(1/max(a$values), digits = 2)
#     camb_katz_wPK <- Rinbix::EpistasisKatz(camb_Adj_partial, alpha, beta = IMP_deg)
#     names(camb_katz_wPK) <- rownames(camb_Adj_partial)
#     camb_katz_wPK.sort <- sort(camb_katz_wPK, T)
#     top_camb_katz_wPK <- camb_katz_wPK.sort[1:200]
#     katz_wPK_inner.features <- c(katz_wPK_inner.features, as.character(names(top_camb_katz_wPK)))
#     katz_wPK_features <- c(katz_wPK_features, as.character(names(top_camb_katz_wPK)))
#   }
#   camb_katz_wPK.expr <- camb_fltr.expr[, unique(katz_wPK_inner.features)]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_katz_wPK.expr), label = camb_class)
#   camb_katz_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=1, eta = .01, nthread = 2,
#                                        nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                        subsample = 0.7, objective = "binary:logistic")
# 
#   camb_katz_wPK_bst.pred <- predict(camb_katz_wPK_bst.model, as.matrix(camb_katz_wPK.expr))
#   katz_wPK_train_acc <- c(katz_wPK_train_acc, accuracy(camb_class, camb_katz_wPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_katz_wPK.expr <- jap_fltr.expr[, unique(katz_wPK_inner.features)]
#   jap_katz_wPK_bst.pred <- predict(camb_katz_wPK_bst.model, as.matrix(jap_katz_wPK.expr))
#   katz_wPK_test_acc <- c(katz_wPK_test_acc, accuracy(jap_class, jap_katz_wPK_bst.pred)$prop.correct)
# }
load("katz_wPK_train_acc.RData")
load("katz_wPK_test_acc.RData")

k_w_train <- katz_wPK_train_acc[which.min(katz_wPK_train_acc - katz_wPK_test_acc)]
k_w_test <- katz_wPK_test_acc[which.min(katz_wPK_train_acc - katz_wPK_test_acc)]
```

### Epistasis-Katz
```{r}
# ekatz_wPK_features <- NULL
# ekatz_wPK_train_acc <- NULL
# ekatz_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ekatz_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- Rinbix::regainInbix(regain_matrix)
#     # ekatz_wPK centrality
#     beta <- diag(camb_Regain$reGAIN)
#     diag(camb_Regain$reGAIN) <- 0
#     camb_ekatz_wPK <- Rinbix::EpistasisKatz(camb_Regain$reGAIN, IMP_deg, beta)
#     names(camb_ekatz_wPK) <- colnames(camb_expr)
#     camb_ekatz_wPK.sort <- sort(camb_ekatz_wPK, T)
#     top_camb_ekatz_wPK <- camb_ekatz_wPK.sort[1:200]
#     ekatz_wPK_inner.features <- c(ekatz_wPK_inner.features, as.character(names(top_camb_ekatz_wPK)))
#     ekatz_wPK_features <- c(ekatz_wPK_features, as.character(names(top_camb_ekatz_wPK)))
#   }
#   camb_ekatz_wPK.expr <- camb_fltr.expr[, unique(ekatz_wPK_inner.features)]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ekatz_wPK.expr), label = camb_class)
#   camb_ekatz_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=2, eta = .1, nthread = 2,
#                                         nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                         subsample = 0.4, objective = "binary:logistic")
# 
#   camb_ekatz_wPK_bst.pred <- predict(camb_ekatz_wPK_bst.model, as.matrix(camb_ekatz_wPK.expr))
#   ekatz_wPK_train_acc <- c(ekatz_wPK_train_acc, accuracy(camb_class, camb_ekatz_wPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_ekatz_wPK.expr <- jap_fltr.expr[, unique(ekatz_wPK_inner.features)]
#   jap_ekatz_wPK_bst.pred <- predict(camb_ekatz_wPK_bst.model, as.matrix(jap_ekatz_wPK.expr))
#   ekatz_wPK_test_acc <- c(ekatz_wPK_test_acc, accuracy(jap_class, jap_ekatz_wPK_bst.pred)$prop.correct)
# }
load("ekatz_wPK_train_acc.RData")
load("ekatz_wPK_test_acc.RData")

ek_w_train <- ekatz_wPK_train_acc[which.min(ekatz_wPK_train_acc - ekatz_wPK_test_acc)]
ek_w_test <- ekatz_wPK_test_acc[which.min(ekatz_wPK_train_acc - ekatz_wPK_test_acc)]
```

### EpistasisRank
```{r}
# ER_wPK_features <- NULL
# ER_wPK_train_acc <- NULL
# ER_wPK_test_acc <- NULL
# # create outer folds
# outer_folds <- caret::createFolds(camb_class, 5, list = FALSE)
# # create inner folds
# for (i in 1:5){
#   ER_wPK_inner.features <- NULL
#   inner_folds <- caret::createFolds(camb_class[outer_folds!=i], 5, list = TRUE)
#   for (j in 1:length(inner_folds)){
#     fold_idx <- which(outer_folds != i)[-inner_folds[[j]]]
#     # reGAIN
#     camb_expr <- camb_fltr.expr[fold_idx,]
#     class <- camb_class[fold_idx]
#     regain_matrix <- data.frame(camb_expr, class)
#     rownames(regain_matrix) <- NULL
#     camb_Regain <- Rinbix::regainInbix(regain_matrix)
#     # ER_wPKrank centrality
#     camb_ER_wPKrank <- Rinbix::EpistasisRank(camb_Regain$reGAIN, IMP_deg)
#     top_camb_ER_wPK <- camb_ER_wPKrank[1:200,]
#     ER_wPK_inner.features <- c(ER_wPK_inner.features, as.character(top_camb_ER_wPK$gene))
#     ER_wPK_features <- c(ER_wPK_features, as.character(top_camb_ER_wPK$gene))
#   }
#   camb_ER_wPK.expr <- camb_fltr.expr[, unique(ER_wPK_inner.features)]
#   # Training
#   camb_dtrain <- xgb.DMatrix(as.matrix(camb_ER_wPK.expr), label = camb_class)
#   camb_ER_wPK_bst.model <- xgb.train(data=camb_dtrain, max.depth=2, eta = .1, nthread = 2,
#                                      nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1,
#                                      subsample = 0.7, objective = "binary:logistic")
# 
#   camb_ER_wPK_bst.pred <- predict(camb_ER_wPK_bst.model, as.matrix(camb_ER_wPK.expr))
#   ER_wPK_train_acc <- c(ER_wPK_train_acc, accuracy(camb_class, camb_ER_wPK_bst.pred)$prop.correct)
# 
#   # Testing
#   jap_ER_wPK.expr <- jap_fltr.expr[, unique(ER_wPK_inner.features)]
#   jap_ER_wPK_bst.pred <- predict(camb_ER_wPK_bst.model, as.matrix(jap_ER_wPK.expr))
#   ER_wPK_test_acc <- c(ER_wPK_test_acc, accuracy(jap_class, jap_ER_wPK_bst.pred)$prop.correct)
# }
load("ER_wPK_train_acc.RData")
load("ER_wPK_test_acc.RData")

er_w_train <- ER_wPK_train_acc[which.min(ER_wPK_train_acc - ER_wPK_test_acc)]
er_w_test <- ER_wPK_test_acc[which.min(ER_wPK_train_acc - ER_wPK_test_acc)]
```

## XGBoost
```{r}
# Training
camb_dtrain <- xgboost::xgb.DMatrix(as.matrix(camb_fltr.expr), label = camb_class)
camb_xgb_model <- xgboost::xgb.train(data=camb_dtrain, max.depth=2, eta = .1, nthread = 2, 
                                   nround=1, gamma = 0, min_child_weight = 1, max_delta_step = 1, 
                                   subsample = 0.4, objective = "binary:logistic")

camb_xgb_pred <- predict(camb_xgb_model, as.matrix(camb_fltr.expr))
xgb_train <- SDMTools::accuracy(camb_class, camb_xgb_pred)$prop.correct
xgb_train

# Testing
jap_xgb_pred <- predict(camb_xgb_model, as.matrix(jap_fltr.expr))
xgb_test <- SDMTools::accuracy(jap_class, jap_xgb_pred)$prop.correct
xgb_test
```


## Accuracy Plots
```{r, fig.width=6, fig.height=6}
PR.Katz_df <- data.frame(
  Data = factor(c("Cambridge-Train", "Japan-Test", "Cambridge-Train", "Japan-Test", "Cambridge-Train", "Japan-Test",
                  "Cambridge-Train", "Japan-Test", "Cambridge-Train", "Japan-Test"), levels = c("Cambridge-Train", "Japan-Test")),
  Method = factor(c("PR", "PR", "Katz", "Katz", "xgb", "xgb", "PR", "PR", "Katz", "Katz"), levels = c("PR", "Katz", "xgb")),
  Accuracy = c(pr_wo_train, pr_wo_test, k_wo_train, k_wo_test, xgb_train, xgb_test, pr_w_train, pr_w_test, k_w_train, k_w_test),
  Status  = c( "no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge",
               "Prior Knowledge", "Prior Knowledge", "Prior Knowledge", "Prior Knowledge")
)

EK.ER_df <- data.frame(
  Data = factor(c("Cambridge-Train", "Japan-Test", "Cambridge-Train", "Japan-Test", "Cambridge-Train", "Japan-Test",
                  "Cambridge-Train", "Japan-Test", "Cambridge-Train", "Japan-Test"), levels = c("Cambridge-Train", "Japan-Test")),
  Method = factor(c("ER", "ER", "EK", "EK", "xgb", "xgb", "ER", "ER", "EK", "EK"),  levels = c("ER", "EK", "xgb")),
  Accuracy = c(er_wo_train, er_wo_test, ek_wo_train, ek_wo_test, xgb_train, xgb_test, er_w_train, er_w_test, ek_w_train, ek_w_test),
  Status = c("no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge", "no Prior Knowledge",
             "Prior Knowledge", "Prior Knowledge", "Prior Knowledge", "Prior Knowledge")
)


p1 <- ggplot2::ggplot(PR.Katz_df, ggplot2::aes(x = Data, y = Accuracy, group = Method)) +
  ggplot2::ggtitle("Co-expression Networks") + ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::geom_point(ggplot2::aes(shape=Method), size = 2) + ggplot2::geom_line(ggplot2::aes(linetype=Method)) +
  ggplot2::theme(axis.title.x=ggplot2::element_blank()) +
  ggplot2::scale_linetype_manual(values=c("twodash", "solid", "dashed"))+
  ggplot2::facet_wrap(~Status) +
  ggplot2::theme(legend.position=c(0.95, 0.5), legend.text=ggplot2::element_text(size=6), legend.title = ggplot2::element_text(size=6))
p2 <- ggplot2::ggplot(EK.ER_df, ggplot2::aes(x = Data, y = Accuracy, group = Method)) +
  ggplot2::ggtitle("Epistasis Networks") + ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::geom_point(ggplot2::aes(shape=Method), size = 2) + ggplot2::geom_line(ggplot2::aes(linetype=Method)) +
  ggplot2::theme(axis.title.x=ggplot2::element_blank()) +
  ggplot2::scale_linetype_manual(values=c("twodash", "solid" ,"dashed"))+
  ggplot2::facet_wrap(~Status) +
  ggplot2::theme(legend.position=c(0.95, 0.5), legend.text=ggplot2::element_text(size=6), legend.title = ggplot2::element_text(size=6))
```


Training accuracy (Cambridge data) and independent validation accuracy (Japan data) with centrality feature selection without prior knowledge (left panels) and with prior knowledge (right panels). Top: co-expression net-work centrality feature selection methods, PageRank (PR) and Katz. Bottom row: expression-epistasis network centrality methods, EpistasisRank (ER) and EpistasisKatz (EK). Accuracies computed by xgboosted trees with nested cross-validation. Xgboost accuracies without feature selection also shown (squares).
```{r}
gridExtra::grid.arrange(p1, p2, nrow = 2)
```





